\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{multirow}

\title{Are large software companies bound to adopt the free software model?}
\author{Merwan Achibet}
\institute{Université du Havre}
\date{}

\definecolor{grey}{RGB}{85,98,112}
\definecolor{white}{RGB}{255,255,255}

\setbeamercolor{background canvas}{bg=grey}
\setbeamercolor{title}{fg=white}
\setbeamercolor{structure}{bg=grey!70, fg=white}
\setbeamercolor{frametitle}{fg=white}
\setbeamercolor{normal text}{bg=grey,fg=white}
\setbeamercolor{block title}{bg=grey!70}
\setbeamercolor{block body}{bg=normal text.bg!90!black}

\useoutertheme{split}
\useinnertheme{rounded}

\setbeamertemplate{items}[ball]
\setbeamertemplate{blocks}[rounded][shadow=true]
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[page number]

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\maketitle

\begin{frame}
  \frametitle{Table of Contents}
  \tableofcontents
\end{frame}

\section{Free software}

\subsection{Definition}

\begin{frame}

  \frametitle{What is free software?}

  \begin{itemize}
    \item<1->{Distributed with its source code}
    \item<2->{Modifiable}
    \item<3->{Redistribuable}
  \end{itemize}

  \pause

  \vfill

  \only<4->{
    \begin{block}{}
      Free software $\neq$ open source
    \end{block}
  }

  \only<5>{
    \begin{block}{}
      Free software $\neq$ Free of charge
    \end{block}
  }

\end{frame}

\subsection{History}

\begin{frame}

  \frametitle{Historically speaking...}

  \begin{block}{1950s}
    \begin{center}Softwares are academic, sharing is the norm.\end{center}
  \end{block}

  \begin{center} $\downarrow$ \end{center}

  \begin{block}{1970s}
    \begin{center}Softwares are commercial products.\end{center}
  \end{block}

  \begin{center} $\downarrow$ \end{center}

  \begin{block}{Near future}
    \begin{center}?\end{center}
  \end{block}

\end{frame}

\section{Software companies and free software}

\subsection{Advantages}

\begin{frame}

  \frametitle{Advantages}

  \begin{columns}

    \column{.2\textwidth}
    \Huge{\textbf{+}}

    \column{.8\textwidth}
    \begin{itemize}
      \item<2->{Reputation}
      \item<3->{Trust}
      \item<4->{No patenting fees}
      \item<5->{No legal fees}
      \item<6->{Built on other free softwares}
      \item<7>{User contribution}
    \end{itemize}

  \end{columns}

\end{frame}

\subsection{Disadvantages}

\begin{frame}

  \frametitle{Disadvantages}

  \begin{columns}

    \column{.2\textwidth}
    \Huge{\textbf{-}}

    \column{.8\textwidth}
    \begin{itemize}
      \item<2->{Intimidating to the user}
      \item<3->{Potentially free $=$ not as good}
      \item<4>{No secret}
    \end{itemize}

  \end{columns}

\end{frame}

\section{Different business models}

\begin{frame}

  \begin{block}{Question}
    Free software as many advantages but is it economically viable?
  \end{block}

  \vfill

  BLAHBLAH

\end{frame}

\subsection{Ubuntu}

\begin{frame}

  \frametitle{Ubuntu Foundation - Ubuntu}

  \begin{block}{Their product}
    A Linux-based operating system.
  \end{block}

  \vfill

  \begin{itemize}
    \item<2->{Deployment}
    \item<3->{Maintenance}
    \item<4>{Training}
  \end{itemize}

\end{frame}

\subsection{Firefox}

\begin{frame}

  \frametitle{Mozilla Foundation - Firefox}

  \begin{block}{Their product}
    A famous internet browser.
  \end{block}

  \vfill

  \begin{itemize}
    \item<2->{Donations}
    \item<3>{Google partnership}
  \end{itemize}

\end{frame}

\subsection{Chrome}

\begin{frame}

  \frametitle{Google - Chrome}

  \begin{block}{Their product}
    An internet browser.
  \end{block}

  \vfill

  \begin{itemize}
    \item<2->{Zero revenue}
    \item<3->{The real product is advertising}
    \item<4>{Chrome lets people access Google more easily}
  \end{itemize}

\end{frame}

\begin{frame}

  \frametitle{Conclusion}

  \begin{block}{}
    \begin{center}Are large software companies bound to adopt the free software
    model?\end{center}
  \end{block}

  YES NO MAYBE

\end{frame}

\end{document}
